---
title: "Analyse d'expansion"
author: "Baptiste Crinière-Boizet"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
---

```{r echo = FALSE,warning=FALSE, message=FALSE, results='hide'}
library(plotly)
library(DT)
library(tidyverse)
setwd("/home/baptiste.criniere/Documents/PB_NC_MB/Data/")
data1 <- openxlsx::read.xlsx("expansion_Barbier.xlsx", sheet = 1)
data2 <- openxlsx::read.xlsx("expansion_Barbier.xlsx", sheet = 2)

data2 <- data2 %>% 
  dplyr::select(-"FID",-"Sex", -"Onset.FTD",-"Onset.ALS", -"ALS.(1/0)", -"age.at.death", -"FTD.disease.duration.at.death",
               -"ALS.disease.duration.at.death" )
data1 <- data1 %>% 
  dplyr::rename("ALS" = "ALS.(1/0)")
gene <- names(data1)[10:45]
outcome <- names(data1)[4:9]
data <- data1 %>% 
  dplyr::left_join(data2, by = "ID") %>% 
  dplyr::mutate(Onset.FTD = Onset.FTD %>% as.numeric) %>% 
  dplyr::mutate(Onset.ALS = Onset.ALS %>% as.numeric) %>% 
  dplyr::mutate(age.at.death = age.at.death %>% as.numeric) %>% 
  dplyr::mutate(FTD.disease.duration.at.death = FTD.disease.duration.at.death %>% as.numeric) %>% 
  dplyr::mutate(ALS.disease.duration.at.death = ALS.disease.duration.at.death %>% as.numeric) %>% 
  dplyr::mutate("ALS.(1/0)" =" ALS.(1/0)" %>% as.numeric)
  
M <- max(as.numeric(data$FID), na.rm = TRUE)
for (i in 1:nrow(data)){
  if (data$FID[i] %in% c(NA, "na")){
    data$FID[i] <- M + i
  }
}
data <- data
# Patient SLA-SAL-BLA-030-008 .
data <- data %>% 
  dplyr::mutate(FID = FID %>% 
                  factor) %>% 
  dplyr::mutate(ALS = ifelse(ALS %in% ".", NA, ALS)) %>% 
  dplyr::mutate(ALS = ALS %>% factor)
```

## Onset FTD {.tabset}
### Table des résultats
```{r echo=FALSE, message=FALSE, warning=FALSE}
result <- data.frame(gene = gene, min.coeff = NA, max.coeff = NA, min.pval = NA, max.pval = NA)
for (i in 1:length(gene)){
  a <- paste0(gene[i],".x")
  b <- paste0(gene[i], ".y")
  d <- paste(a, by = "+", b)
  formula <- paste(outcome[1], by = " ~ Sex + (1|FID) +", d)
  model <- lmerTest::lmer(formula = formula, data = data) 
  result$min.coeff[i] <- coef(summary(model))[3,1]
  result$max.coeff[i] <- coef(summary(model))[4,1]
  result$min.pval[i] <- coef(summary(model))[3,5]
  result$max.pval[i] <- coef(summary(model))[4,5]
}
result <- result %>% 
 mutate_if(is.numeric, round, digits = 4)
DT::datatable(result, rownames = FALSE, colnames = c("Gene", "Min coeff", "Max coeff", "Min pvalue", "Maxpvalue"))
```

### Manhanttan plot {.tabset}
#### Min
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(min.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

#### Max
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(max.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

## Onset ALS {.tabset}
### Table des résultats
```{r echo=FALSE, message=FALSE, warning=FALSE}
result <- data.frame(gene = gene, min.coeff = NA, max.coeff = NA, min.pval = NA, max.pval = NA)
for (i in 1:length(gene)){
  if (!gene[i] %in% c("PABPN1")){
    a <- paste0(gene[i],".x")
    b <- paste0(gene[i], ".y")
    d <- paste(a, by = "+", b)
    formula <- paste(outcome[2], by = " ~ Sex + (1|FID) +", d)
    model <- lmerTest::lmer(formula = formula, data = data) 
    result$min.coeff[i] <- coef(summary(model))[3,1]
    result$max.coeff[i] <- coef(summary(model))[4,1]
    result$min.pval[i] <- coef(summary(model))[3,5]
    result$max.pval[i] <- coef(summary(model))[4,5]
  }
}
result <- result %>% 
 mutate_if(is.numeric, round, digits = 4)
DT::datatable(result, rownames = FALSE, colnames = c("Gene", "Min coeff", "Max coeff", "Min pvalue", "Maxpvalue"))
```

### Manhanttan plot {.tabset}
#### Min
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(min.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

#### Max
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(max.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

## Age at death {.tabset}
### Table des résultats
```{r echo=FALSE, message=FALSE, warning=FALSE}
result <- data.frame(gene = gene, min.coeff = NA, max.coeff = NA, min.pval = NA, max.pval = NA)
for (i in 1:length(gene)){
  if (!gene[i] %in% c("PABPN1")){
    a <- paste0(gene[i],".x")
    b <- paste0(gene[i], ".y")
    d <- paste(a, by = "+", b)
    formula <- paste(outcome[4], by = " ~ Sex + (1|FID) +", d)
    model <- lmerTest::lmer(formula = formula, data = data) 
    result$min.coeff[i] <- coef(summary(model))[3,1]
    result$max.coeff[i] <- coef(summary(model))[4,1]
    result$min.pval[i] <- coef(summary(model))[3,5]
    result$max.pval[i] <- coef(summary(model))[4,5]
  }
}
result <- result %>% 
 mutate_if(is.numeric, round, digits = 4)
DT::datatable(result, rownames = FALSE, colnames = c("Gene", "Min coeff", "Max coeff", "Min pvalue", "Maxpvalue"))
```

### Manhanttan plot {.tabset}
#### Min
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(min.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

#### Max
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(max.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

## FTD disease duration at death {.tabset}
### Table descriptive
```{r echo=FALSE, message=FALSE, warning=FALSE}
result <- data.frame(gene = gene, min.coeff = NA, max.coeff = NA, min.pval = NA, max.pval = NA)
for (i in 1:length(gene)){
  if (!gene[i] %in% c("PABPN1")){
    a <- paste0(gene[i],".x")
    b <- paste0(gene[i], ".y")
    d <- paste(a, by = "+", b)
    formula <- paste(outcome[5], by = " ~ Sex + (1|FID) +", d)
    model <- lmerTest::lmer(formula = formula, data = data) 
    result$min.coeff[i] <- coef(summary(model))[3,1]
    result$max.coeff[i] <- coef(summary(model))[4,1]
    result$min.pval[i] <- coef(summary(model))[3,5]
    result$max.pval[i] <- coef(summary(model))[4,5]
  }
}
result <- result %>% 
 mutate_if(is.numeric, round, digits = 4)
DT::datatable(result, rownames = FALSE, colnames = c("Gene", "Min coeff", "Max coeff", "Min pvalue", "Maxpvalue"))
```

### Manhanttan plot {.tabset}
#### Min
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(min.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

#### Max
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(max.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```




## ALS disease duration at death {.tabset}
### Table descriptive
```{r echo=FALSE, message=FALSE, warning=FALSE}
result <- data.frame(gene = gene, min.coeff = NA, max.coeff = NA, min.pval = NA, max.pval = NA)
for (i in 1:length(gene)){
  if (!gene[i] %in% c("PABPN1")){
    a <- paste0(gene[i],".x")
    b <- paste0(gene[i], ".y")
    d <- paste(a, by = "+", b)
    formula <- paste(outcome[6], by = " ~ Sex + (1|FID) +", d)
    model <- lmerTest::lmer(formula = formula, data = data) 
    result$min.coeff[i] <- coef(summary(model))[3,1]
    result$max.coeff[i] <- coef(summary(model))[4,1]
    result$min.pval[i] <- coef(summary(model))[3,5]
    result$max.pval[i] <- coef(summary(model))[4,5]
  }
}
result <- result %>% 
 mutate_if(is.numeric, round, digits = 4)
DT::datatable(result, rownames = FALSE, colnames = c("Gene", "Min coeff", "Max coeff", "Min pvalue", "Maxpvalue"))
```

### Manhanttan plot {.tabset}
#### Min
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(min.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

#### Max
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(max.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```




## ALS {.tabset}
### Table descriptive
```{r echo=FALSE, message=FALSE, warning=FALSE}
result <- data.frame(gene = gene, min.coeff = NA, max.coeff = NA, min.pval = NA, max.pval = NA)
for (i in 1:length(gene)){
    a <- paste0(gene[i],".x")
    b <- paste0(gene[i], ".y")
    d <- paste(a, by = "+", b)
    formula <- paste(outcome[3], by = " ~ Sex + (1|FID) +", d)
    model <- lme4::glmer(formula = formula, data = data, family = binomial) 
    result$min.coeff[i] <- coef(summary(model))[3,1]
    result$max.coeff[i] <- coef(summary(model))[4,1]
    result$min.pval[i] <- coef(summary(model))[3,4]
    result$max.pval[i] <- coef(summary(model))[4,4]
    if (!is.null(summary(model)$optinfo$conv$lme4$messages)){
      result[i, 2:5] <- NA
    }
}
result <- result %>% 
 mutate_if(is.numeric, round, digits = 4)
DT::datatable(result, rownames = FALSE, colnames = c("Gene", "Min coeff", "Max coeff", "Min pvalue", "Maxpvalue"))


#str_detect(summary(model)$optinfo$conv$lme4$messages, "Model failed to converge") & 
```

### Manhanttan plot {.tabset}
#### Min
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(min.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```

#### Max
```{r echo=FALSE, message=FALSE, warning=FALSE}
result$id <- 1:nrow(result)
fig <- result %>% 
  ggplot(aes(x = id, y = -1*log10(max.pval), label = gene))+
  geom_point(color = "dodgerblue3")+
  geom_hline(yintercept = 1.3, color = "darkorange1",linetype = "dashed")+
  geom_hline(yintercept = 2.85, color = "darkorange3", linetype = "dashed")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())+
  labs(y = "-log10(p)")+
  ylim(c(0,4))
  
fig <- ggplotly(tooltip = c("label"))
htmltools::div(fig,align="center")
```


